6CCS3PRJ \\\vspace{0.5cm} A Pragmatic Infrastructure Modelling and Verification Framework

### 2014-04-14 paper.tex
Removed implementation specifics from design goals

\paragraph{Simple Install Process} The install process is a one-line bash command that downloads and executes the bootstrap shell script, that in turn detects system architecture and downloads the appropriate binary, copying it into the system path and configuring it to run as a service.

When the binary is first executed, it performs node registration by making a PUT request to the monitoring tool’s REST API with the hostname of the server, and supplementary meta-data.

\textbf{Example Install Command:}
\begin{minted}[fontfamily=fi4]{bash}
  curl -s https://management-server/agent/install | sh
\end{minted}

\paragraph{Automatic Service Detection} The tool builds in a basic service detection process that is executed during node registration and at rare periodic intervals after this point.
 It detects the presence of known services such as Apache2, Nginx or MySQL within the currently executing process table, these services are defined in a simple component definition list bundled with the agent software, making it easy to expand in future. It then takes further steps to extract version numbers from installed services and includes this information in the node meta-data.

There are a number of issues that prevent this process from being accurate in every scenario, but in practice it is a flexible solution that can detect most services accurately enough to be a useful time saving feature.

@@

@@

\pagebreak
\section{System Architecture}
\label{section:architecture}

The nature of the system is inherently distributed. Service specifications are written by a developer on their personal machines, must be executed on any number of target servers and have results returned to a centralised location. The following decomposition of components is suggested:

@@

The API itself could assume this role, but the number of events recieved could become very high and overwhelm the API. When considering that a very large percentage of events will be identical to previous events, it makes sense only for the significant events to be forwarded to the API.

Due to its highly parallel and asynchronous nature, this service will be written in Clojure, as the functional programming style and asynchronous primitives within the language make it an ideal candidate for this type of processing. The Riemann ~\cite{Riemann} event aggregation and stream processing library has been chosen for stream buffering and filtering.

@@

    component "Nginx" do
      description "The web server"
    end
  end
\end{minted}
\label{fig:MetaProgrammingSnippet}
\pagebreak

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this. The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message, which may be printed to the console or serialised and sent to the event stream.

@@

Importantly, as the blocks executed are pure Ruby code that can throw exceptions, any arbitrary code that fits these criterion can be used within a block:

@@

\section{Monitoring Agent}

@@

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\@. It provides fast feedback similar to running a suite of unit tests and can be used to quickly validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarily used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependencies in PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialise each object to JSON, it is then sent as a PUT request to the management API\. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Pull] - Pulls an existing service from the management server - This command will download the original service specification file from a previous `push" command. This can be used to conveniently add new hosts to existing services.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process, which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

\subsection{Service Discovery}

@@

>>>>

@@

Based on decomposition of the problem, the following concepts are deemed to be relevant:

@@

\pagebreak
\section{System Architecture}

\label{section:architecture}

The nature of the system is inherently distributed. Service specifications are written by a developer on their personal machines, must be executed on any number of target servers and have results returned to a centralised location. The following decomposition of components is suggested, with futher explanation provided below:

@@

The API itself could assume this role, but the number of events recieved could become very high and overwhelm it. When considering that a very large percentage of events will be identical to previous events, it makes sense only for the significant events to be forwarded to the API.

Due to its highly parallel and asynchronous nature, this service will be written in Clojure, as the functional programming style and asynchronous primitives within the language make it an ideal candidate for this type of processing. The Riemann ~\cite{Riemann} event aggregation and stream processing library has been chosen for stream buffering and filtering, the rationale for this choice is expanded upon in the next section.

@@

    component "Nginx" do
      description "The web server"
    end
  end
\end{minted}
\label{fig:MetaProgrammingSnippet}

The most important feature of the DSL is the ability to encode checks as Ruby blocks. A block contains one or more assertions that throw exceptions with descriptive error messages if they fail. Below is a simple example of this:

@@

The health check will be executed every 30 seconds by the monitoring agent. The monitoring agent will catch any exceptions and return a descriptive error message, which may be printed to the console or serialised and sent to the event stream.

Importantly, as the blocks executed are pure Ruby code that can throw exceptions, any arbitrary code that fits these criterion can be used within a block, including standard expressions and the use of external libraries as plugins:

@@

\pagebreak
\section{Monitoring Agent}

@@

\subsection{Workflow}

After installation, a service can be specified and monitored in the following way:

\begin{description}
  \item[Creation or Discovery] The tool is used to create a service specification, either manually or using the service discovery feature.
  \item[Local Testing] The specification is checked for validity on the local machine, ensuring that checks are valid and test for the desired properties. Iterative improvements can be made.
  \item[Pushing] The service definition is pushed to the configured management server, where it is used to update the data model and therefore the user interface. The file is interpreted and posted to the API as a series of partial updates. The original file is also sent, so comments and formatting are maintained for future reference.
  \item[Pulling] Where other hosts provide the same service, they can ``pull" the service configuration from the management server, taking a copy of the original file. They will then be notified of subsequent pushes, ensuring the specifications stay in sync. If this is not appropriate, this role can be fulfilled by an application deployment or configuration management process.
  \item[Starting] Monitoring is started on the host, from this point onwards, checks will be executed as appropriate and reported back to the management server. Unexpectedly terminating this process will lead to incident alerts.
\end{description}

@@

\begin{description}
  \item[Check] - Run configured health checks on demand. This command is typically used on developer workstations as they iteratively build a service specification using the DSL\@. It provides fast feedback similar to running a suite of unit tests and can be used to quickly validate assumptions, before pushing the specification to the management server.
  \item[Discover] - Creates a service specification from detected system services. This feature can be used primarily to reverse-engineer existing infrastructure setups into a service specification. More details about the discovery mechanism are provided in the next subsection.
  \item[Graph] - Render a graph of the current service - This can be primarily used on the developer workstation, it will use the GraphViz library to render a graph of the system and its dependencies in PDF format. This can be used on its own as a form of documentation, but is typically intended to help visualise the service specification and its dependencies to ensure its correctness.
  \item[Push] - Pushes the specified service to a management server - This command will parse the DSL and serialise each object to JSON, it is then sent as a PUT request to the management API\@. Changes to the service specification take effect immediately and update any connected users in near real-time.
  \item[Pull] - Pulls an existing service from the management server - This command will download the original service specification file from a previous `push" command. This can be used to conveniently add new hosts to existing services.
  \item[Setup] - Register this host with the management server - This is used on the servers running the system under test, it will associate a host with a defined service, listing it in the Web UI and accepting events and check data from it.
  \item[Start] - Starts a monitoring daemon process, which runs in the background and executes checks according to their defined intervals. It will report all data back to the management server's event stream endpoint. Any errors occurring during this process will be logged to a file for later analysis.
  \item[Stop] - Stop monitoring the specified service.
  \item[Restart] - Restart monitoring of the specified service.
  \item[Help] - Describes the function of all available commands.
\end{description}

\pagebreak
\subsection{Service Discovery}

@@

\subsection{Bundled Checks \& Extensibility}

The tool bundles a number of built in checks for common use cases, they can be used by calling methods of the same name, such as:

\begin{itemize}
  \item {Running} - Confirms that a process is running. A process can quit unexpectedly for a number of reasons including misconfiguration, resource constraints, and application errors.
  \item {Listening} - Confirms that a process is listening on the specified network port.
  \item {HTTP Success} - Confirms that a HTTP endpoint responds with a successful status code, that is, one that falls within the 200-300 range when redirects are followed.
  \item {CPU Load} - Uses the systems 5 minute load average to determine CPU utilization, scaled to the number of CPUs. A high figure can suggest that the current CPU specification is insufficient to support the application.
  \item {Process CPU Load} - Returns the current CPU utilization for a particular process, useful in identifying processes that are behaving incorrectly.
  \item {Memory Usage} - Returns the total amount of free memory, this metric is sometimes less informative due to aggressive caching and memory management behavior of modern Linux kernels.
  \item {Process Memory Usage} - Returns the total amount of memory used by a particular process, useful in identifying processes that are behaving incorrectly.
  \item {Disk Usage} - Returns the disk usage specifics for a given mount point. Useful for identifying overflowing log files or underspecified hardware resources.
\end{itemize}
